---
title: "VAR Estimation"
author: "Scheer et al, 2022"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
knitr::opts_chunk$set(
  eval = T,
  echo = T,
  message = F,
  warning = F
)
```

```{r}
rm(list=ls())
library(vars)
library(urca)
library(tidyverse)
library(here)
library(plotly)
library(kableExtra)

select <- dplyr::select
```

# Data and Functions

Load the relevant data

```{r}
df <- read_csv(here("data","use","oxmetrics.csv"),na = ".", col_t = "Ddddd")
```


Load the functions used in this estimation

```{r}
source(here("code/functions_VAR.R"))
```

Data wrangling

```{r}
order = c("WTI", "emp_total", "discount", "total_oil")
series <- ts(df[-1],start = 1976,frequency = 12)
series <- na.trim(series)
series <- series[,order]
plot(series)
```

# Lag Selection

Select the correct lag

```{r}
lagselection <- VARselect(series,type = "both",season = 12)

lagselection$criteria %>% 
  t() %>% 
  as_tibble() %>%
  mutate(Lag = 1:10) %>%
  relocate(Lag) %>% 
  setNames(gsub("(n)","",names(.),fixed=T)) %>% 
  kable(digits = 3,
        caption = "Lag Selection in a Trend and Intercept Specification with Annual Seasonality") %>% 
  kable_styling(full_width = FALSE) %>% 
  kable_paper()
```

# Estimation

Use the Johansen Procedure for estimating the VAR

```{r}
cajo <- ca.jo(series, type = "eigen", ecdet = "trend", K = 2, season = 12,spec = "transitory")
summary(cajo)
```

Output a table:

```{r}
cajo@cval %>% 
  as_tibble() %>% 
  setNames(gsub("pct","% CV",names(.))) %>% 
  mutate(Rank = c("$r \\leq 3$","$r \\leq 2$","$r \\leq 1$","$r = 0$"),
         `Test Stat.`= cajo@teststat) %>% 
  relocate(Rank, `Test Stat.`) %>% 
  kable(digits = 2, escape = FALSE, caption = "Johansen Test with 2 Lags, a Trend Specification and Annual Seasonality") %>% 
  kable_paper() %>% 
  kable_styling(full_width = FALSE) 
```

# Restricted Vector Error Correction Model

```{r}
rrr <- cajorls(cajo,r = 2)

kable(rrr$rlm$coefficients,digits = 3,caption = "Restricted Vector Error Correction Model using OLS") %>% 
  kable_paper() %>% 
  kable_styling(full_width = FALSE)
```

Show the cointegrating relationships

```{r}
kable(rrr$beta,digits = 3,caption = "Co-Integrating Relationships") %>% 
  kable_paper() %>% 
  kable_styling(full_width = FALSE)
```
# Testing for Exogeneity of Alpha (OxMetrics)


We test for exogeneity of alpha in OxMetrics 8. Here we include the OxMetrics 8 Code needed to confirm our results in the SM: 

```{Ox, eval = FALSE}
// Batch Code
module("PcGive");
package("PcGive", "Multiple-equation");
usedata("oxmetrics.csv");
system
{
    Y = emp_total, WTI, discount, total_oil;
    Z = emp_total_1, emp_total_2, WTI_1, WTI_2, discount_1, discount_2, 
        total_oil_1, total_oil_2, Trend;
    U = Constant, Seasonal, Seasonal_1, Seasonal_2, Seasonal_3, 
        Seasonal_4, Seasonal_5, Seasonal_6, Seasonal_7, Seasonal_8, 
        Seasonal_9, Seasonal_10;
}
rank(2);
constraints
{
&9=1;
&10=0;
&15=1;
&14=0;



}
estimate("COINT", 2007-03-01, 0, 2020-09-01, 0);
```


```{Ox, eval = FALSE}
// Batch code for SYS(13)
module("PcGive");
package("PcGive", "Multiple-equation");
usedata("oxmetrics.csv");
system
{
    Y = emp_total, WTI, discount, total_oil;
    Z = emp_total_1, emp_total_2, WTI_1, WTI_2, discount_1, discount_2, 
        total_oil_1, total_oil_2, Trend;
    U = Constant, Seasonal, Seasonal_1, Seasonal_2, Seasonal_3, 
        Seasonal_4, Seasonal_5, Seasonal_6, Seasonal_7, Seasonal_8, 
        Seasonal_9, Seasonal_10;
}
rank(2);
constraints
{
&9=1;
&10=0;
&15=1;
&14=0;

&2=0;
&3=0;


}
estimate("COINT", 2007-03-01, 0, 2020-09-01, 0);

```



# Robustness Tests

Variables to determine:

-   Season yes or no (`season = 12`)
-   `VARselect_type`: options are const, trend and both
-   trace or eigen for the ca.jo test in `cajotesttype`
-   `spec` which can be longrun or transitory
-   `ecdet` which decides on the inclusion of the intercept or the trend in the cointegration, "none", "const" or "trend"

After considering all possible models, it's clear that results are robust to using either 'trace' or 'eigen' for the ca.jo test in `cajotesttype`. We therefore choose `trace`. Similarly, `spec` (which can be longrun or transitory) does not influence results. We therefore choose `longrun`. Even `VARselect_type` is robust, where options are const, trend and both. To formulate the widest possible model, we choose `both`.


```{r}
var_list <- list()
for(seas in c(0,12)){
  for(seltype in c("both")) {
    for(cajotype in c("trace")){
      for(cajospec in c("longrun")){
        for(cajoecdet in c("none","const","trend")){
          print(paste(seas,seltype,cajotype,cajospec, cajoecdet,sep = "-"))
          
          run <- vecm_fun(df = df,
                          name = paste(seas,seltype,cajotype,cajospec, cajoecdet,sep = "-"),
                          season = seas,
                          VARselect_type = seltype,
                          spec = cajospec,
                          ecdet = cajoecdet,
                          cajotesttype = cajotype,
                          cajotestpval = "1pct")
          
          var_list[[paste(seas,seltype,cajotype,cajospec, cajoecdet,sep = "_")]] <- run
        }
      }
    }
  }
}
```

# Impulse Response Functions

```{r}
subset_var <- names(var_list)

extract_irf(vecm_list = var_list[subset_var],
            impulse = "WTI",variable = "emp_total") %>% 
  group_by(model) %>% 
  mutate(x = 1:n()) %>% 
  ggplot(aes(x=x,y=emp_total,color=model)) + 
  geom_line() -> plot

ggplotly(plot)
```


# Plotting

```{r}
extract_irf(vecm_list = var_list,
            impulse = "WTI",
            variable = "emp_total") %>% 
  group_by(model) %>% 
  mutate(x = 1:n()) %>% 
  ungroup %>% 
  separate(col = model,sep = "-",
           into = c("Seasonality","varselect","cajotest","cajospec","VECM Specification"),
           remove = FALSE) %>% 
  mutate(`VECM Specification` = case_when(`VECM Specification`=="none"~"No Intercept or Trend",
                                          `VECM Specification`=="const"~"Intercept",
                                          TRUE~stringr::str_to_title(`VECM Specification`)),
         `VECM Specification` = factor(`VECM Specification`,levels = c("No Intercept or Trend","Intercept","Trend")),
         Seasonality = case_when(Seasonality==0~"No Seasonality",
                                 Seasonality==12~"Annual Seasonality")) %>% 
  
  
  
  
  ggplot(aes(x=x,y=emp_total,color=model)) + 
  geom_hline(aes(yintercept = 0)) +
  geom_ribbon(aes(ymin = emp_total_lower,ymax = emp_total_upper,fill=model),alpha = 0.3)+
  geom_line(size = 1) + 
  
  facet_grid(Seasonality~`VECM Specification`)+
  
  scale_x_continuous(breaks = c(seq(2,12,2)))+
  scale_y_continuous(labels = function(x){scales::comma(x*1000)})+
  
  scale_color_viridis_d(name = "Model")+
  scale_fill_viridis_d(name = NULL, guide = FALSE)+
  
  labs(title = "Impulse Response Function",
       subtitle = "US$1 Impulse to WTI",
       x = "Months after Impulse",
       y = "Total Employment") +
  
  theme(legend.position = "none",
        strip.background = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(colour = "grey"),
        legend.key = element_rect(fill = NA))

```

```{r}
extract_irf(vecm_list = var_list,
            impulse = "discount",
            variable = "emp_total") %>% 
  group_by(model) %>% 
  mutate(x = 1:n()) %>% 
  ungroup %>% 
  separate(col = model,sep = "-",
           into = c("Seasonality","varselect","cajotest","cajospec","VECM Specification"),
           remove = FALSE) %>% 
  mutate(`VECM Specification` = case_when(`VECM Specification`=="none"~"No Intercept or Trend",
                                          `VECM Specification`=="const"~"Intercept",
                                          TRUE~stringr::str_to_title(`VECM Specification`)),
         `VECM Specification` = factor(`VECM Specification`,levels = c("No Intercept or Trend","Intercept","Trend")),
         Seasonality = case_when(Seasonality==0~"No Seasonality",
                                 Seasonality==12~"Annual Seasonality")) %>% 
  
  
  
  
  ggplot(aes(x=x,y=emp_total,color=model)) + 
  geom_hline(aes(yintercept = 0)) +
  geom_ribbon(aes(ymin = emp_total_lower,ymax = emp_total_upper,fill=model),alpha = 0.3)+
  geom_line(size = 1) + 
  
  facet_grid(Seasonality~`VECM Specification`)+
  
  scale_x_continuous(breaks = c(seq(2,12,2)))+
  scale_y_continuous(labels = function(x){scales::comma(x*1000)})+
  
  scale_color_viridis_d(name = "Model")+
  scale_fill_viridis_d(name = NULL, guide = FALSE)+
  
  labs(title = "Impulse Response Function",
       subtitle = "US$1 Impulse to the WTI - WCS Discount",
       x = "Months after Impulse",
       y = "Total Employment") +
  
  theme(legend.position = "none",
        strip.background = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(colour = "grey"),
        legend.key = element_rect(fill = NA)) 

```

```{r}
extract_irf(vecm_list = var_list,
            impulse = "total_oil",
            variable = "emp_total") %>% 
  group_by(model) %>% 
  mutate(x = 1:n()) %>% 
  ungroup %>% 
  separate(col = model,sep = "-",
           into = c("Seasonality","varselect","cajotest","cajospec","VECM Specification"),
           remove = FALSE) %>% 
  mutate(`VECM Specification` = case_when(`VECM Specification`=="none"~"No Intercept or Trend",
                                          `VECM Specification`=="const"~"Intercept",
                                          TRUE~stringr::str_to_title(`VECM Specification`)),
         `VECM Specification` = factor(`VECM Specification`,levels = c("No Intercept or Trend","Intercept","Trend")),
         Seasonality = case_when(Seasonality==0~"No Seasonality",
                                 Seasonality==12~"Annual Seasonality")) %>% 
  
  
  
  
  ggplot(aes(x=x,y=emp_total,color=model)) + 
  geom_hline(aes(yintercept = 0)) +
  geom_ribbon(aes(ymin = emp_total_lower,ymax = emp_total_upper,fill=model),alpha = 0.3)+
  geom_line(size = 1) + 
  
  facet_grid(Seasonality~`VECM Specification`)+
  
  scale_x_continuous(breaks = c(seq(2,12,2)))+
  scale_y_continuous(labels = function(x){scales::comma(x*1000)})+
  
  scale_color_viridis_d(name = "Model")+
  scale_fill_viridis_d(name = NULL, guide = FALSE)+
  
  labs(title = "Impulse Response Function",
       subtitle = "1M Impulse to Oil Production",
       x = "Months after Impulse",
       y = "Total Employment") +
  
  theme(legend.position = "none",
        strip.background = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(colour = "grey"),
        legend.key = element_rect(fill = NA)) 
```
